### 2017-08-15 README.Rmd
adds data package note
expand description and add references

## Description

`methyvim` is an R package that provides facilities for differential methylation
analysis based on _variable importance measures_ (VIMs), a class of statistical
target parameters that arise in causal inference.

@@

1. The average treatment effect (ATE) for discrete exposures/treatments,
2. A nonparametric variable importance measure (NPVI) for continuous exposures
[@chambaz2012estimation].

@@

1. Pre-screening of genomic sites is used to isolate a subset of sites for
which there is evidence of differential methylation. Targeted minimum
loss-based estimates of VIMs are computed for this subset of sites only (for
the sake of computational feasibility). Several screening approaches are
available, adapting core routines from the following R packages:
[`limma`](http://bioconductor.org/packages/release/bioc/html/limma.html),
[`tmle.npvi`](https://CRAN.R-project.org/package=tmle.npvi),
[`adaptest`](https://github.com/nhejazi/adaptest).
2. Nonparametric VIMs are estimated for the specified parameter, currently using
routines from the R packages
[`tmle.npvi`](https://CRAN.R-project.org/package=tmle.npvi) to estimate the
NPVI parameter, or [`tmle`](https://CRAN.R-project.org/package=tmle) to
estimate the ATE.
3. Since pre-screening is performed prior to estimating VIMs, we make use of a
multiple testing correction uniquely suited to such settings, specifically we
use the modified marginal Benjamini & Hochberg step-up False Discovery Rate
controlling procedure for multi-stage analyses (FDR-MSA) [@tuglus2009modified].

See @vdl2011targeted for a general description of the framework of targeted
minimum loss-based estimation.

@@

>>>>

## What's `methyvim`?

`methyvim` is an R package that provides facilities for differential methylation
analysis based on _variable importance measures_ (VIMs), a class of statistical
estimable target parameters that arise in causal inference.

@@

1. The _average treatment effect_ (ATE) for discrete exposures/treatments,
2. A _nonparametric variable importance measure_ (NPVI) for continuous exposures
[@chambaz2012estimation].

@@

1. Pre-screening of genomic sites is used to isolate a subset of sites for
which there is cursory evidence of differential methylation. For the sake of
computational feasibility, targeted minimum loss-based estimates of VIMs are
computed only for this subset of sites. Several screening approaches are
available, adapting core routines from the following R packages:
[`limma`](http://bioconductor.org/packages/release/bioc/html/limma.html),
[`tmle.npvi`](https://CRAN.R-project.org/package=tmle.npvi).
2. Nonparametric VIMs are estimated for the specified parameter, currently using
routines from the R packages
[`tmle.npvi`](https://CRAN.R-project.org/package=tmle.npvi) (to estimate the
NPVI parameter) or [`tmle`](https://CRAN.R-project.org/package=tmle) (to
estimate the ATE).
3. Since pre-screening is performed prior to estimating VIMs, we make use of a
multiple testing correction uniquely suited to such settings. Rather generally,
the Benjamini & Hochberg procedure for controlling the False Discovery Rate
(FDR) [@benjamini1995controlling] is applied due to multiple testing issues
inherent to the estimation procedure. Specifically, we apply a modified marginal
Benjamini & Hochberg step-up False Discovery Rate controlling procedure for
multi-stage analyses (FDR-MSA) [@tuglus2009modified].

For a general discussion of the framework of targeted minimum loss-based
estimation, the many applications of this methodology, and the role the
framework plays in statistical causal inference, the recommended references are
@vdl2011targeted and @vdl2017targeted. @hernan2018causal and @pearl2009causality
may be of interest to those desiring a more general introduction to statistical
causal inference.

@@

## Related

* [R/`methyvimData`](https://github.com/nhejazi/methyvimData) - R package with
    sample experimental DNA methylation data for use as an example with this
    analysis package.
